---
title: "**Cheatsheets Links**"
author: "Gabriel I. Cook"
#date: "`r Sys.Date()`"
date: "`r format(Sys.time(), '%d %B, %Y')`"

execute:
  #enabled: false
  freeze: auto

order: 15
---

```{r message=FALSE, warning=FALSE, include=FALSE}
restart_r_session <- FALSE   # restart R in case of undefined objects  
run_commonR <- FALSE         # run the _common.R script
source(here::here("src", "module_content", "module_readings_and_videos.R"))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
no_proceed_to_install <- FALSE
under_construction    <- FALSE # print a warning message
source(here::here("src", "module_content", "module_call_outs.R"))
```



```{r}
# Load necessary libraries
library(dplyr)
library(knitr)

# Define a function to get all PDF links
get_all_pdf_links <- function(url) {
  library(rvest)
  
  # Read the webpage
  page <- try(read_html(url), silent = TRUE)
  
  # Handle the case where the URL isn't a valid webpage
  if (inherits(page, "try-error")) {
    return("Unable to read the webpage. Please check the URL.")
  }
  
  # Extract all links on the page
  links <- page |> html_nodes("a") |> html_attr("href")
  
  # Filter for valid URLs (ignoring 'null' values)
  valid_links <- links[!is.na(links) & links != ""]

  # Create full URLs (GitHub links)
  full_links <- paste0("https://github.com", valid_links)

  # Remove duplicates
  full_links <- unique(full_links)
  
  # Filter for PDF links
  pdf_links <- full_links[grepl("\\.pdf$", full_links, ignore.case = TRUE)]

  return(pdf_links)
}

# Define a function to create a data frame for the table
create_pdf_table <- function(pdf_links) {
  data.frame(
    Name = basename(pdf_links) |> tools::file_path_sans_ext(),
    URL = pdf_links,
    Cheatsheet = basename(pdf_links) |> tools::file_path_sans_ext(),
    stringsAsFactors = FALSE
  )
}

# Example usage
url <- "https://github.com/rstudio/cheatsheets"
all_pdf_links <- get_all_pdf_links(url)
pdf_table <- create_pdf_table(all_pdf_links)

# Print the table
kable(pdf_table, caption = "Cheatsheets PDF Links")
